---
title: "ST661 R prog Assignment 2"
author: "Paul Williamson 18145469"
date: "6 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=4, fig.height=4,fig.align='center', prompt=FALSE,
                      comment=NA)
```

#QUESTION 1:
  
_Download the cdc1.Rdata file from moodle and load_
```{r}
load("C:/Users/PAUL/Documents/.Maynooth HDip/ST661 R programming/Assignment2/cdc1.Rdata")
```

_check data imported_
```{r}
str(cdc1)
head(cdc1)
```

*Q1(a) Change variables exerany, smoke100 and hlthplan to be factors with suitable levels.*

_Use a char vector to index cols into data_
```{r}
cols_to_change <- c("exerany","smoke100","hlthplan")
cdc1[cols_to_change] <- lapply(cdc1[cols_to_change], factor)
```

_check class types in cdc1 dataset to confirm variables have been changed to factors_
```{r}
sapply(cdc1, class)
```

_Alternatively change each col separately using..._
_cdc$exerany <-as.factor(cdc$exerany)_

_change to suitable levels_
```{r}
levels(cdc1$exerany) <- c("no", "yes")
levels(cdc1$hlthplan) <- c("nocover", "cover")
levels(cdc1$smoke100) <- c("nosmoke", "smoke")
```

_Confirm changes to levels_
```{r}
str(cdc1)
head(cdc1)
```

*Q1(b) Calculate the proportion of men that have health very good or better. Do the same for women.*
_Who has the better health?_
```{r}
class(cdc1)

hlth_table <- table(cdc1$genhlth, cdc1$gender)
class(hlth_table)
hlth_table
barplot(hlth_table, beside=TRUE, col=1:5, legend.text=TRUE)
```

_Compare overall proportions on barplot_
```{r}
prop.table(hlth_table)
barplot(prop.table(hlth_table), col=1:5, legend.text=TRUE)
```

_get proportions per gender_
```{r}
male_total <-sum(hlth_table[,"m"]) # sum of all rows in column 1 (m)
male_total
male_vg_ex_prop <- sum(hlth_table[1:2,"m"])/male_total # proportion: sum/total
male_vg_ex_prop

female_total <-sum(hlth_table[,"f"]) # sum of all rows in column 2 (f)
female_total
female_vg_ex_prop <- sum(hlth_table[1:2,"f"])/female_total # proportion: sum/total
female_vg_ex_prop
```

*Q1(c) Use subset to extract the smokers.*
_For the smokers, calculate the proportion of men that have health very good or better._
_Do the same for women._

```{r}
head(cdc1)
smokers_table <- subset(cdc1, smoke100 == "smoke")
head(smokers_table)
```

_extract male smokers_
```{r}
healthy_male_smokers <- subset(smokers_table, gender == "m" & (genhlth == "excellent" | genhlth == "very good"))
head(healthy_male_smokers)
healthy_male_smokers <- subset(healthy_male_smokers, select = c(genhlth,smoke100,gender))
head(healthy_male_smokers)

proportion_healthy_male_smokers <- length(healthy_male_smokers)/male_total
proportion_healthy_male_smokers
```

_extract female smokers_
```{r}
healthy_female_smokers <- subset(smokers_table, gender == "f" & (genhlth == "excellent" | genhlth == "very good"))
head(healthy_female_smokers)
healthy_female_smokers <- subset(healthy_female_smokers, select = c(genhlth,smoke100,gender))
healthy_female_smokers

proportion_healthy_female_smokers <- length(healthy_female_smokers)/female_total
proportion_healthy_female_smokers
```

*(d) Repeat (c) for non smokers.*
```{r}
non_smokers_table <- subset(cdc1, smoke100 == "nosmoke")
head(non_smokers_table)
```

_extract male non smokers_
```{r}
healthy_male_non_smokers <- subset(non_smokers_table, gender == "m" & (genhlth == "excellent" | genhlth == "very good"))
head(healthy_male_non_smokers)
healthy_male_non_smokers <- subset(healthy_male_non_smokers, select = c(genhlth,smoke100,gender))
head(healthy_male_non_smokers)

proportion_healthy_male_non_smokers <- length(healthy_male_non_smokers)/male_total
proportion_healthy_male_non_smokers
```

_extract female smokers_
```{r}
healthy_female_non_smokers <- subset(non_smokers_table, gender == "f" & (genhlth == "excellent" | genhlth == "very good"))
head(healthy_female_non_smokers)
healthy_female_non_smokers <- subset(healthy_female_non_smokers, select = c(genhlth,smoke100,gender))
healthy_female_non_smokers

proportion_healthy_female_non_smokers <- length(healthy_female_non_smokers)/female_total
proportion_healthy_female_non_smokers
```

*(e) Based on your calculations of (c) and (d), compare the health of men and women, and the*
    *health of smokers and non smokers.*
```{r}
proportion_healthy_male_smokers
proportion_healthy_female_smokers
proportion_healthy_male_non_smokers
proportion_healthy_female_non_smokers
```

#QUESTION 2:

*Q2(a) Write function called skew to calculate the skewness of a numeric vector.*

_m3 function definition_
```{r}
m3 <- function(vect_of_samples, vect_mean = mean(vect_of_samples),no_samples = length(vect_of_samples),answer = 0){
  vect_mean <- sum(vect_of_samples)/length(vect_of_samples)
  no_samples <- length(vect_of_samples)
  for(i in vect_of_samples){
  answer <- answer + (i - vect_mean)^3/no_samples
  }
  answer
  }
```

_m2 function definition_
```{r}
m2 <- function(vect_of_samples, vect_mean = mean(vect_of_samples),no_samples = length(vect_of_samples),answer = 0){
  vect_mean <- sum(vect_of_samples)/length(vect_of_samples)
  no_samples <- length(vect_of_samples)
  for(i in vect_of_samples){
    answer <- answer + (i - vect_mean)^2/no_samples
  }
  answer
}
```

_Skew function definition_
```{r}
skew <- function(vect_of_samples,answer = 0){
  answer <- m3(vect_of_samples)/(m2(vect_of_samples)^(3/2))
  answer
}
```

*Q2(b) Test your function on mtcars$wt*
```{r}
skew(mtcars$wt)
```

*Q2(c) Test it also on a set of 50 numbers generated from the standard normal distribution.*
```{r}
rand_50_norm_distr <- rnorm(50,mean=0,sd=1)
rand_50_norm_distr
skew(rand_50_norm_distr)
```

*Q2(d) Using replicate, replicate the calculation of (c) 1000 times. Draw a histogram of the results*
*and calculate the mean skewness value.*
```{r}
replicate_normal_1000 <- replicate(1000,rnorm(50, mean=0, sd=1))
hist(replicate_normal_1000)
skew(replicate_normal_1000)
```

*Q2(e) Repeat (c) and (d), this time using a set 50 numbers generated from the*
*exponential distribution with parameter 1.*

_50 numbers generated from the exponential distribution_
```{r}
replicate_expo_1000 <- replicate(1000,rexp(50, rate = 1))
hist(replicate_expo_1000)
skew(replicate_expo_1000)
```

*Q2(f) Compare the results of (d) and (e)*

*Q2(g) Write code which calculates the skewness for every numeric variable in cdc1.*
_Your code should construct a named vector of skewness values, with NA's for non numeric variables._
_Your code should not give any warnings._
```{r}
class(cdc1)
str(cdc1)

skewness_vect <- vector("numeric", length=length(cdc1))
skewness_vect

for (i in 1:length(cdc1)){
  if (is.numeric(cdc1[[i]]))
    skewness_vect[i] <- skew(cdc1[[i]])
  else skewness_vect[i]<- NA
}
skewness_vect
```

#QUESTION 3:

*Q3(a) Make a plot of weight-wtdesire versus age for the cdc1 data. Use pch=20 and a colour vector*
      *which assigns colours as follows:*

_  weight > wtdesire and age > 30 red_
_  weight > wtdesire and age <= 30 blue_
_  weight <= wtdesire and age > 30 orange_
_  weight <= wtdesire and age <= 30 cyan_
_  Hint: use ifelse._
```{r}
plot(weight ~ wtdesire, data=cdc1, col= ifelse(weight>wtdesire & cdc1$age>30,"red",
                                        ifelse(weight>wtdesire & cdc1$age<=30,"blue",
                                        ifelse(weight<wtdesire & cdc1$age>30, "orange",
                                        ifelse(weight<wtdesire & cdc1$age<=30,"cyan","black")))),
                                        pch=20)
```

*Q3(b) Recall in a previous lecture we used boxplot.stats to find indices of outliers. Write a function*
*using boxplot.stats called boxplot.out that given a numeric vector returns low" if there are*
*low outliers, "high" if there are high outliers, "both" if there are high and low outliers, and*
*"none" if there are no outliers.*
```{r}
boxplot.out <- function(input_vector){
  
  check_outliers_status <- "."
  boxplot_stats_vector <- boxplot.stats(input_vector)$'stats'
  
  check_low_outlier <- boxplot_stats_vector[1]
  temp_LQ <- boxplot_stats_vector[2]
  temp_median <- boxplot_stats_vector[3]
  temp_UQ <- boxplot_stats_vector[4]
  temp_IQR <- temp_UQ - temp_LQ
  check_high_outlier <- boxplot_stats_vector[5]

  ifelse((check_low_outlier < (temp_LQ - (1.5*temp_IQR))) & (check_high_outlier > (temp_UQ + (1.5*temp_IQR))),
         check_outliers_status <- "both",
         ifelse(check_high_outlier > (temp_UQ + (1.5*temp_IQR)),
         check_outliers_status <- "high",
         ifelse(check_low_outlier < (temp_LQ - (1.5*temp_IQR)),
         check_outliers_status <- "low",
         check_outliers_status <- "none"
         )))
  
  check_outliers_status
}
```

_test boxplot.out function using cdc1$weight as input vector_
```{r}
boxplot.out(cdc1$weight)
```

*Q3(c) For the cdc1 data, make boxplots showing weight-wtdesire versus genhlth.*
```{r}
plot(weight ~ wtdesire, data=cdc1, col= ifelse(cdc1$genhlth == "excellent","red",
                                        ifelse(cdc1$genhlth == "very good","blue",
                                        ifelse(cdc1$genhlth == "good", "orange",
                                        ifelse(cdc1$genhlth == "fair","cyan","black")))),
                                        pch=20)

boxplot_collection <- boxplot(weight~wtdesire, data = cdc1, col=cdc1$genhlth)
str(boxplot_collection)
#boxplot_collection
```

*Q3(d) Using boxplot.out and tapply, construct a vector indicating the type of outlier*
      *(high, low, both or none) in each boxplot.*
```{r}
outlier_vector <- vector("numeric",length(cdc1))
for (i in 1:length(cdc1)){
  if (is.numeric(cdc1[[i]]))
    outlier_vector[i] <- boxplot.out(cdc1[[i]])
  else outlier_vector[i]<- "NA"
}
outlier_vector
```

*Q3(e) Construct a color vector that is "red" for boxplots that have high and low outliers, "blue"*
*for just high outliers, "green" for just low outliers, and "yellow" otherwise. Use this vector to*
*colour the boxplots.*
```{r}
colours_vector <- c("red", "blue", "green", "yellow")
```

#QUESTION 4:
```{r}
load("C:/Users/PAUL/Documents/.Maynooth HDip/ST661 R programming/Assignment2/kobe.RData")
head(kobe)
```
*Q4(a) Write a while loop to calculate how many throws are required to reach 3 hoops.*
```{r}
scores <- 0
num_games <- length(kobe$basket)
count_shots <- 0
while(scores<3){
  count_shots <- count_shots +1
  if(kobe$basket[count_shots]=="H"){scores <- scores + 1}
  }
count_shots
```

*Q4(b) Write another while loop to calculate how many throws are required to reach 3 more hoops*
```{r}
while(scores<6){
  # re-using the count_shots (=5) from previous snippet
  count_shots <- count_shots +1
  if(kobe$basket[count_shots]=="H"){scores <- scores + 1}
}
count_shots
```


# FINAL EXAM - QUESTION 1:

**(A) TREE DIAGRAM**

                            /-----98%---[+ive] = 1.96%
               /---[lupus]--
              2%            \------2%---[-ive] = 0.04%
             /                
      [test]
             \
              98%            /----26%---[+ive] = 25.48%
               \---[no lupus]
                             \----74%---[-ive] = 72.52%

                          
**(B) PROB OF TESTING POSITIVE WITH LUPUS:**

  [LUPAS & +IVE] = 0.02 * 0.98 = 0.0196 = 1.96%
  
  [NO LUPUS & +IVE] = 0.98 * 0.26 = 0.2548 = 25.48%
  
  [+IVE] = 0.0196 + 0.2548 = 0.2744 = 27.44%
  
  => (0.0196 / 0.2744) = 0.0714 = 7.1%

  
**(C) 'DR HOUSE' IS RIGHT (1 - 0.0714) = 0.9286 => 92.86% OF THE TIME!**
