---
title: "ST661 Assignment1"
author: "Paul Williamson, 18145469"
date: "12/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

**(a) Construct a vector with values 10,11,12,13,14.**
_Do this using :, c and seq. Assign your result to x._
```{r}
x1 <- (10:14)
x1

x2 <- c(10,11,12,13,14)
x2

x3 <- seq(10, 14) 
x3
```


**(b) Construct a vector with values 14,10,11,13,12. Call it y.**
```{r}
y <- c(14,10,11,13,12)
y
```


**(c) Extract the third value of x.**
```{r}
x1[3]
```


**(d) Extract the second and fourth values of y.**
```{r}
y[2]; y[4]
```

**(e) Write a sentence explaining what the next line of code does.**
```{r}
result <- sum(x1>y)
result
message("This function makes an element-wise comparison of the elements in vector x\nwith the elements in vector y and count those elements in vector x which are larger than\ntheir positional counterpart in vector y; In this case the result is 3.")
```

**(f) Write a sentence explaining what the next line of code does. (Look up the help page for abs.)**
```{r}
result <- sum(abs(x2-y) == 1)
result
message("The (x-y) returns a 'difference' vector with values equal to the element-wise difference between vector x values and vector y values;
the abs() returns a vector with the absolute (positive) magnitude of the values in that 'difference' vector;
the '== 1' returns a 'logical' vector with values true for those values equal to 1 and false for any other value not equal to 1;
the sum function results in a summation of all the elements in the resulting 'logical' vector, it treats a false value as 0 and a true value as 1;
the overall effect or result is a count of the number of elements in vector x which are of magnitude 1 away from their positional counterpart in vector y.")
```

**(g) Write code which implements the following formula. summation[i=1:5]Xi^2/Yi**
```{r}
x <- 1:5
y <- x
result <- sum((x*x)/y)
result
```


**(h) Form a 5 x 2 matrix using the values in x and y.**
```{r}
new_matrix <- cbind(x3,y)
new_matrix
```


# Question 2

```{r}
house <- read.csv("C:/Users/PAUL/Documents/.Maynooth HDip/ST661 R programming/Assignment1/house.csv",row.names=1)
head(house)
```

**(a) Use head to check the first few rows of the data.**
```{r}
head(house)
```

**(b) Extract the row of the data for Ireland.**
```{r}
house["Ireland",]
```

**(c) Find Spain's value for TooDark.**
```{r}
house["Spain",1]
```

**(d) For how many countries is the percentage without an indoor toilet 0?**
```{r}
count_NoIndoorToilet <- sum(house[,4] == 0)
count_NoIndoorToilet
```

**(e) For how many countries is the percentage without an indoor toilet and without a bath/shower both 0?**
```{r}
count_NoIndoorToilet_NoBathShower <- sum((house[,4] == 0) & (house[,3] == 0))
count_NoIndoorToilet_NoBathShower
```

**(f) Find the names of the countries where the percentage without an indoor toilet and without a**
**bar/shower are both zero. Your code should give a vector of the country names.**
```{r}
no_modCons <- subset(house, (house[,4] == 0) & (house[,3] == 0))
no_modCons <- rownames(no_modCons)
no_modCons
```

**(g) Find the maximum value of the variable Damp.**
```{r}
max_damp <- max(house$Damp)
max_damp
```


# Question 3

```{r}
load("C:/Users/PAUL/Documents/.Maynooth HDip/ST661 R programming/Assignment1/worldcup.Rdata")
head(gtimes) # display data
head(ngoals) # display data
```

**(a) Draw a barplot of the number of goals per match. What is the most frequent number of goals?**
**Find the mean and the median.**

```{r}
barplot(ngoals) # plot

mode_goals <- max(unique(ngoals))

mean_goals <- mean(ngoals)
mean_goals

median_goals <- median(ngoals)
median_goals
```


**(b) The first 52 matches were played in 1990, the next 52 matches were played in 1994, the next 64**
**matches were played in 1998, and the last 64 matches were played in 2002. Therefore elements**
**52, 104, 168 and 232 of ngoals give the numbers of matches scored in the four finals. Extract**
**these elements of ngoals into a vector.**

```{r}
goalsScoredInFinals <- ngoals[c(52,104,168,232)]
goalsScoredInFinals
```

**(c) Make a vector of length 232 containing the year each match was played. Call your vector year.**
**(Hint: use rep.)**
```{r}
yr1990 <- 52
yr1994 <- 52
yr1998 <- 64
yr2002 <- 64
year <- c(rep(1990, yr1990),rep(1994, yr1994),rep(1998, yr1998),rep(2002, yr2002))
year
```

**(d) Turn the vector year into a factor. Make a data frame where the first variable is ngoals and**
**the second is year.**
```{r}
year <- factor(year)
year

year_goals_df <- data.frame(ngoals,year)
year_goals_df
```

**(e) Make a vector called matchn containing the number of the match for that world cup. The**
**vector should be 1..52, 1..52, 1..64, 1..64. Add this vector to the data frame.**
```{r}
matchn <- c(seq(1,52),seq(1,52),seq(1,64),seq(1,64))
matchn
year_goals_df <- cbind(year_goals_df,matchn)
year_goals_df
```


**(f) Extract the rows 52, 104, 168 and 232 of the data frame. Which year's final has the most goals?**
```{r}
max_FinalsGoals <- max(year_goals_df[c(52,104,168,232),1])
max_FinalsGoals
```

**(g) Make a boxplot of the number of goals played for each year. What can you say about the 8 goal match?**
```{r}
boxplot(year_goals_df$ngoals)
message("What can you say about the 8 goal match? It would seem to be an outlier.")
```

**(h) (Advanced, optional). The vector gtimes contains the cumulative time in minutes from the**
**start of the 1990 world cup in which goals were scored (counting playing time only.) Construct**
**another vector minute containing the minute of each game in which the goal is scored. In what**
**minutes of the matches are goals most likely to occur?**
```{r}
head(gtimes)
minute <- vector()
currTime <- 1

for (currTime in 1:length(gtimes))
  {
    currentTime <- gtimes[currTime]
    while(currentTime>90)
    {
      currentTime <- currentTime-90
    }
    minute[currTime] <- currentTime
  }
minute

minute <- factor(minute)
table(minute)
names(which.max(table(minute)))
```
