---
title: "ST661 R prog assignment 4"
author: "Paul Williamson, 18145469"
date: "1 December 2018"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
```

```{r libraries, message=FALSE, echo = FALSE}
#install.packages("tidyverse")
library(tidyverse) # every time
```


<!-- For adding pagebreaks to html output   -->
<!-- <P style="page-break-before: always">  -->


## Question 1

```{r load h4data}
load("h4data")
head(wim,20)
```

<P style="page-break-before: always">

>(a) Use mutate (from dplyr) to add a new variable which is the difference in heights of the winner
>    and loser. Use ggplot to draw a histogram of the new variable.

```{r create ht_diff}
wim1 <- mutate(wim, ht_diff = winner_ht - loser_ht)
wim1$ht_diff # confirm new variable ht_diff

ht_diff_hist_plotty <- ggplot(wim1, aes(ht_diff), na.omit()) + geom_histogram()
suppressWarnings(print(ht_diff_hist_plotty))
```

<P style="page-break-before: always">

>(b) Construct a dataset **by_player** with one row per player, recording also the number of wins, ranking points,
>    height and country of the player. Do this in the following steps (i - iv):

i. Construct a dataset w_wim (use **select** and **mutate**), containing for each match the information:
     * name = winner_name
     * points = winner_rank_points
     * ht = winner_ht
     * ioc= winner_ioc
     * A new variable **wins** with a value of **1**
```{r create w_wim}
w_wim <- select(wim1, name=winner_name, points=winner_rank_points, ht= winner_ht, ioc= winner_ioc)
w_wim <- mutate(w_wim, wins=1)
str(w_wim)
```
  
***
  
ii. Construct a dataset l_wim (use **select** and **mutate**), containing for each match the information:
     * name = loser_name
     * points = loser_rank_points
     * ht = loser_ht
     * ioc= loser_ioc
     * A new variable **wins** with a value of **0**
```{r create l_wim}
l_wim <- select(wim1, name=loser_name, points=loser_rank_points, ht= loser_ht, ioc= loser_ioc)
l_wim <- mutate(l_wim, wins=0)
str(l_wim)
```

<P style="page-break-before: always">

iii. Use **rbind** to stick w_wim and l_wim together. Call the result **wl_wim**
```{r combine w_wim & l_wim}
wl_wim <- rbind(w_wim, l_wim)
wl_wim
```
  
***
  
iv. Use **group_by** on wl_wim, to group it by name and **summarise** the result of the previous step with:
     * wins=sum(wins)
     * points=points[1]
     * ht=ht[1]
     * ioc=ioc[1]
```{r use group_by & summarise}
# used pipes to pass input/output arguments between group_by and summarise functions
by_player <- wl_wim %>% group_by(name) %>% summarise(wins=sum(wins), points=points[1], ht=ht[1], ioc=ioc[1])
head(by_player,3)
tail(by_player,3)
```

<P style="page-break-before: always">

>(c) Calculate the average height for all players in the tournament. Use ggplot to plot player points
>    versus number of wins.

```{r avg height}
avg_ht <- mean(by_player$ht,na.rm = TRUE)
avg_ht <- round(avg_ht,2); avg_ht
```

```{r plot wins v points}
pts_wins_plotty <- ggplot(by_player, aes(wins, points)) + geom_point()
pts_wins_plotty

# label points above certain value - not used
# pts_wins_plotty + geom_text(aes(label=ifelse(ht>=avg_ht, ht,'')),hjust=0,vjust=0)
```
  
<P style="page-break-before: always">

>(d) Using the dataset **by_player**, write code to find the names of the tournament winner and the
>    losing finalist. If you did not manage to correctly construct the dataset by_player, do this
>    some other way.

**_This solution using the by_player dataset created above assumes a symmetrical draw..._**
```{r names of winner & losing finalist}
winner_name <- by_player[which(by_player$wins == max(by_player$wins)),]$name;
winner_name

losing_finalist_name <- by_player[which(by_player$wins == max(by_player$wins)-1),]$name;
losing_finalist_name
```
  
***

**_This solution using the original imported dataset assumes the order of the matches is first to last..._**  
     * _Used only to confirm the previous result_
```{r names of winner & losing finalist other}
tail(select(wim, winner_name, loser_name, score),1)
```
  
***
  
>(e) Calculate the number of wins per country. How many matches were won by Spanish (ESP)
>    players?

```{r wins per country & spanish wins}
wins_per_country <- by_player %>% group_by(ioc) %>% summarise(total_wins_per_country = sum(wins))

esp_wins <- wins_per_country[which(wins_per_country$ioc == "ESP"),]$total_wins_per_country;
esp_wins
```
  
<P style="page-break-before: always">

>(f) Draw a barplot showing the number of wins for the top 10 countries, preferrably in decreasing
>    order by wins.

```{r barplot wins per country}
# minus sign sorts descending order
wins_per_country_sorted <- wins_per_country[order(-wins_per_country$total_wins_per_country),];
top_10_countries <- head(wins_per_country_sorted,10);
t(top_10_countries) # transposed just for display purposes

top_10_countries_plotty <- ggplot(top_10_countries, aes(x = reorder(ioc, -total_wins_per_country), 
                                      y = total_wins_per_country, fill = ioc)) +
  geom_col(colour="black") + geom_text(aes(label = total_wins_per_country), hjust=0.5, vjust=-0.5) +
  theme(legend.position="none"); top_10_countries_plotty
```

<P style="page-break-before: always">

## Question 2

**For the University ranking data from the midterm, first convert appropriate variables to numeric.**
**Then answer these questions using dplyr tools and ggplot, in all cases except part (b).**
  
>(a) Summarise the Location (countries) of the universities with counts. Using ggplot, draw a
>    barplot of these location counts.
  
```{r import data}
d <- read.csv("https://tinyurl.com/y8e8cb9t", stringsAsFactors = FALSE);
str(d); # check structure
d$QualityÂ.of.Education <- as.numeric(d$QualityÂ.of.Education)
str(d); # confirm conversion of appropriate variables to numeric
```

***
  
```{r per country}  
univs_per_country <- count(d, Location); univs_per_country;
```

<P style="page-break-before: always">

```{r plot per country}
univs_per_ctry_plotty <- ggplot(univs_per_country, aes(x = Location, y = n)) + geom_col();
univs_per_ctry_plotty
```

<P style="page-break-before: always">

>(b) Redraw the barplot, with bars in decreasing order. Use fct_reorder.

```{r sort plot per country decreasing}
univs_per_ctry_plotty_decreasing <- 
  ggplot(univs_per_country, aes(x = fct_reorder(Location, -n), y = n)) +
  geom_col();

univs_per_ctry_plotty_decreasing
```

***
  
>(c) Write a function called top5 that counts the number of values in 1, 2, ...5 in a numeric vector,
>    ignoring NAs. Use %in% .

```{r top5 function}
top5 <- function(input_vector){
  counter <- 0;
  for (i in 1:length(input_vector)){
    if (input_vector[i] %in% 1:5){
      counter <- counter + 1;
    }
  }
  counter
}
```

<P style="page-break-before: always">

**_Check that the top5 function gives the correct answer on vectors below:_**  
      * x1 <- c(NA,10:1)  
      * x2 <- rnorm(8)  
      * x3 <- c(6,2,8,9,-1,0,4)  
     
```{r check top5}
x1 <- c(NA,10:1); top5(x1);
x2 <- rnorm(8); top5(x2);
x3 <- c(6,2,8,9,-1,0,4); top5(x3);
```   


>(d) Use top5 to calculate for each university, how many of the categories Quality.of.Education, Publications,
>    Influence, Citations, Broad.Impact accorded them a top 5 ranking. Add this information to your dataset.
>      Use base R for your answer.

```{r check top5 univs}
QOE_col <- which(colnames(d)=="QualityÂ.of.Education") # get col number
IMP_col <- which(colnames(d)=="Broad.Impact") # get col number
for (i in 1:nrow(d)){
  d$top5count[i] <- top5(d[i, QOE_col:IMP_col]) # create new col and set value
}
d[1:10, c(which(colnames(d)=="Institution"),ncol(d))] # check result from above
```


>(e) Here is a dplyr answer to the previous question. Explain how it works. Write code which
>    confirms the columns tc and topcount are identical. Here the base R solution is simpler, in
>    my opinion. Can anyone supply a simpler and better dplyr-based answer?

```{r}
sel <- c("QualityÂ.of.Education","Publications","Influence","Citations","Broad.Impact")
str(sel)
d1 <- d %>% mutate_at(sel, '<=', 5) %>% select(one_of(sel)) %>% mutate(tc = rowSums(., na.rm=F))
d1[1:10, ncol(d1)]
```
**
**d1 <- d %>% mutate_at(sel, '<=', 5) %>% select(one_of(sel)) %>% mutate(tc = rowSums(., na.rm=T))**
  
**d1 <- d %>% **  
     - assigns output to dataframe object d1, 'pipes' copy of dataframe object into proceeding function  

**mutate_at(sel, '<=', 5...**
     - uses temp 'sel' to identify/hold cols satisfying condition '<=' 5  

**select(one_of(sel)) %>%**  
     - passes cols previously identified to next function without specifying actual col names  

**mutate(tc =...**  
     - create new col in input dataframe ('tc') to hold sum of row val  

**tc = rowSums(., na.rm=...**  
     - calculate sum of all  ('.') passed row vals, also ignores NAs  

```{r alternative dplyr check top5 univs}
#d2 <- for(i in 1:nrow(d)){ mutate(d, tc2[i] <- sum(d[i, 5:9])) }#d2[1:5,"tc2"]
```


>(f) Construct a dataset containing the names of Universities who have at least one top 5 ranking. -->


>(g) Add a new variable Continent to d, that is NorthAmerica for Universities in USA and Canada,
>    Asia for Japan and South Korea, and Europe otherwise. Suggestion: use ifelse within mutate.

```{r add continent}
d <- mutate(d, Continent = ifelse((d$Location == "USA")|(d$Location =="Canada"), "NorthAmerica", ifelse((d$Location == "Japan")|(d$Location=="South Korea"),"Asia", "Europe" )));
d[1:5, c(which(colnames(d)=="Institution"),ncol(d))]
```

>(h) Plot the Citations versus the Broad.Impact rankings, using different colours to distinguish the
>    continents.
  
```{r}
# using 'Influence' to control point size, minus sign to invert i.e. 1 = most, 40 = least
univs_cita_bro_plotty <- ggplot(d, aes(Citations, Broad.Impact, colour = Continent)) + 
  geom_jitter(aes(size = -Influence)) + guides(size=FALSE); univs_cita_bro_plotty +
  scale_y_reverse(lim=c(400,0)) + scale_x_reverse(lim=c(700,0))


# less cluttered vis.
# univs_cita_bro_plotty <- ggplot(d, aes(sqrt(Citations), Broad.Impact, colour = Continent)) +
#   geom_jitter(aes(size = -Influence)) + guides(size=FALSE); univs_cita_bro_plotty + 
#   scale_y_reverse(lim=c(400,0))
```