---
title: 'ST661 R prog Assignment #3'
author: "Paul Williamson, 18145469"
date: "16 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Question 1

_This code loads the data set Pima.tr, and its help file. Use ggplot2 for all plots._
```{r}
library(ggplot2)
library(MASS)
head(Pima.tr)
str(Pima.tr)
```

_Using this dataset:_
  
a. Make a scatterplot plot of bp versus npreg.
```{r}
plotty <- ggplot(Pima.tr,aes(bp,npreg))
plotty + geom_point()

```

b. Using the function cut_interval, construct a factor version of npreg with n=4 levels.
    + Call this new variable npregf.
    + Add this variable to dataset Pima.tr.
```{r}
npregf <- Pima.tr$npreg
cut_interval(npregf,4)

#head(npregf)
#head(Pima.tr)
#Pima.tr <- cbind(Pima.tr,npregf)
#str(Pima.tr)
#Pima.tr$npregf <- factor(Pima.tr$npregf)
#levels(Pima.tr$npregf)
#levels(Pima.tr$npregf)

```


c. Plot boxplots of bp for each npregf level.


d. Make a scatterplot of glu versus age. Use colour to show variable type and add smooths for the two groups.

e. Redo the previous plot, separating out the two types. (Colour is not now needed)



## Question 2

_The Irish road safety authority recently produced a report with the graph below. Reconstruct it using ggplot. (Use google for help!)_

```{r}
Agegroup <- c("0-15","16-25","26-35","36-45","46-55","56-65","66+")
y16 <- c(10,40,38,23,18,13,44)
y17 <- c(4,33,28,22,22,16,33)
mydata <- data.frame(Agegroup,y16,y17)
""
Year_vect <- c("y16","y17")


#rownames(mydata) <- Agegroup
#colnames(mydata)<-Year_vect
str(mydata)
#mydata1 <-t(mydata)
str(mydata)
#ggplot(data=mydata)+geom_bar(mapping=aes(x=Agegroup))


mydata1 <- mydata.long
  
mydata.long <- gather(mydata, variable, value, -Agegroup)


ggplot(data = mydata, mapping = aes(Year_vect)) + geom_bar(stat="identity", position = "dodge")

```

```{r}
mydata2 <- seq(0:66)
cut(mydata2, 7)

mydata3 <- cbind(Agegroup, c(10,40,38,23,18,13,44),c(4,33,28,22,22,16,33))
mydata4 <- ("2016", )


ggplot(data = mydata, mapping = aes(Year_vect)) + geom_bar(stat="identity", position = "dodge")

```



```{r}
Unit <- c("A", "B", "C", "D")
Yes <- c(50, 65, 20, 41)
No <- c(70, 67, 40, 20)
Missing <- c(10, 12, 8, 7)
df <- data.frame(Unit, Yes, No, Missing)

ggplot(data = df, aes(x = Unit, y = Yes)) +
  geom_col() +
  geom_text(aes(label = Yes), position = position_stack(vjust = 0.5))





ggplot(data = df.long, aes(x = Unit, y = value, fill = variable)) +
  geom_col(position = position_dodge()) 


```


ggplot(means.long,aes(x=variable,y=value,fill=factor(gender)))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_discrete(name="Gender",
                      breaks=c(1, 2),
                      labels=c("Male", "Female"))+
  xlab("Beverage")+ylab("Mean Percentage")


















